\LARGE \bf Approximately Orchestrated Road Traffic Automata\\ a General Traffic Simulation Framework

### 2012-03-13 papers/itsc_2012/final.tex
we have a results section finally!

% \long\def\commentp#1{}
% \long\def\commentpk#1{}
% \long\def\commentm#1{}
% \long\def\commentd#1{}

@@

\pix{map_model.png}
    {Roads contain several lanes, and intersections contain several turns
     between lanes.}
    {scale=0.5}

@@

However, with very high volumes of traffic, it is possible for waiting agents to
accumulate and fill lanes to their capacity. When this happens in a circular
manner so that the agent at the front of each lane is waiting to perform a turn
into another full lane with the same such head agent, gridlock \cite{gridlock}
occurs; no agent in the system will make progress unless an agent at the front
decides to pick a different turn. At best, AORTA can detect this by maintaining
a dependency graph that encodes the ``is following'' relation and finding
cycles. It is not clear how to prevent or mend deadlock without forcing
agents to re-route.

\commentd{is the above clear and an acceptable limitation? should i find other
sources that have the same troubles?}

@@

mention all the fixed nuances, like pause time at stop sign

  stop sign: FCFS, judged by stopping for the right time within the right dist of
  the end of the intersection.

  traffic signal: duration matters, turns grouped together, how timings are
  propagated.

  autonomous reservation: the greedy approach is a bit wretched. either time
  out, or somehow count how many people are on lanes in the queued groups.
  ideally we'd even recurse and factor in

\commentd{upgrading stop signs to signals, optimizing timing?}

>>>>

%\long\def\commentp#1{}
%\long\def\commentpk#1{}
%\long\def\commentm#1{}
%\long\def\commentd#1{}

@@

%\pix{map_model.png}
%    {Roads contain several lanes, and intersections contain several turns
%     between lanes.}
%    {scale=0.5}

@@

% TODO in fact, this is caused by lack of lane changing!!!
% but that's scary... bad routes really do cause it.

However, with very high volumes of traffic, it is possible for waiting agents to
accumulate and fill lanes to their capacity. When this happens in a circular
manner so that the agent at the front of each lane is waiting to perform a turn
into another full lane with the same such head agent, gridlock \cite{gridlock}
occurs; no agent in the system will make progress unless an agent at the front
decides to pick a different turn. AORTA can detect this by finding cycles in the
graph of the ``is following'' relation.

In many observed cases, the gridlock is caused by AORTA's present lack of
lane-changing support: agents loop around these intersections in order to enter
a lane adjacent to their original lane. Since these routes are still legal, it
is desirable to prevent or mend this without forcing different routes. The
conditions under which every vehicle eventually may proceed are known.
\cite{AAAI11-au}.

@@

A particular focus of AORTA is intersection policies. To demonstrate how much
policy affects agent performance, three policies are evaluated: stop sign,
traffic signal, and autonomous reservation. These are compared with all other
parameters fixed:

\begin{enumerate}
  \item the map (a $6$ by $7$ mile slice of downtown Austin, Texas
        \footnote{bounded by the Mopac and Springdale Road horizontally,
        and Oltorf Street and Koenig Lane vertically})
  \item $dt = 0.1$ seconds
  \item the agents spawned and their routes -- a continuous generator creates
        one new agent every $1$ simulation second, with a uniformly random
        starting position and goal
  \item the duration of the experiment: one hour
\end{enumerate}

\subsection{Experiment parameters}

Each policy has parameters, so the configuration used is relevant. Stop signs
accept agents FCFS, refusing agents until they have idled at $0$ speed for $1$
second. This delay approximates human reaction time. The traffic signal policy
requires timings and groups of simultaneous turns for each intersection. The
configuration tested here uses a breadth-first search from one major road's
lanes to try to synchronize green signals for the agents near that lane to not
stop for several intersections. This heuristic needs improvement in how it
handles the remaining groups of turns that are scheduled after the first
sequence. Finally, the autonomous reservation policy allows new agents to join
existing groups of agents if their turn is compatible with the others. To
prevent one heavy direction of traffic from hogging the intersection
indefinitely, a timeout preempts and cycles through reservations. Ideally this
should be based on the relative numbers of agents depending on their
reservation.

\subsection{Results}

\pix{avg_lag_agent_atx.png}
    {The distribution of average delay reveals that traffic signals can cause the most lag.}
    {scale=0.35}

If an intersection immediately permits an agent to proceed and grants other
agents in the same lane similar access (because this new agent must still follow
these others), then the new agent may simply proceed at full speed limit without
pausing. This gives the least possible delay spent along some lane, so any extra
time is defined as lag and is used to measure intersection performance. On
average, reservation policies cause less than $1$ second of delay at each
intersection, stop signs cause about $7$ seconds lag, and traffic signals cause
$16$ seconds.

\pix{agent_cnt_atx.png}
    {Autonomous reservations outperform stop signs, and traffic signal gridlocked.}
    {scale=0.35}

The number of agents changes because the continuous generator introduces a new
agent every second. This count reaches a steady state when the generator
introduces an agent at the same rate as older agents finish their route. As one
measure of intersection policy efficiency, it reveals that autonomous
reservations are twice as efficient as the competition, with a steady-state of
about $400$ agents. The count for traffic signals continues to increase because
gridlock occurred in one segment of the map, causing agents to not finish their
route. Unexpected starvation was also observed: although the signals cycle
through turns regularly, agents cannot proceed when the previous cycle caused a
destination lane to fill and become blocked.

Each policy has its trade-offs in terms of performance and physical hardware
cost, suggesting a mixed approach, perhaps with intersection policies adapting
as traffic conditions shift. Researchers could conduct this experiment and many
others without needing to understand most of AORTA's code-base. The largest
intersection policy, traffic signals with the flooding heuristic for timing
assignment, is only $400$ lines of Scala.
